\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-07-05 archive.tex
algorithm flowchart

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
%\usetikzlibrary{decorations.pathmorphing}
\usepackage{dsfont}
\usepackage{amsmath}

@@

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
\end{figure}
\begin{figure}[h]
\centering
%  \includegraphics{figures/iterations.eps}
\begin{tikzpicture}
  \begin{scope}[shift={(2.5,0)}]
    \node[obs] at (3,3) (dx){$\mathcal{D}$};
    \node[latent] at (3.5,1.75) (zp){$p$};
%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
    \node[obs] at (3.5,0.7) (xp) {$x$};
    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
%    \draw [->,red] (tx) -- (zp);
    \draw [->] (dx) -- (zp);
    \node at (nn) {
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
  \end{scope}

>>>>

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration=snake}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
  \caption{\system solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
\end{figure}

\begin{figure}
  \begin{tikzpicture}
    \node at (0,0) (d){DSL};
    \node at ([yshift = -2cm]d) (t){$\text{Task}$};

    \node at ([xshift = 2cm]t) (nn){
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[align = center, text width = 1cm] at ([yshift = 0.6cm]nn.north) {\baselineskip=0pt \small Recognition model\par};
    \draw [->] (t.east) -- ([xshift = -0.5cm]nn.west);

    \node[draw,rounded corners, inner sep = 10] at ([xshift = 4.5cm,yshift = -1cm]) (s){Search};

    \draw [->] ([xshift = 0.5cm]nn.east) -- ([yshift = -0.25cm]s.west);
    \draw [->,rounded corners] (d.east) -- ([yshift = 2cm]nn.center) -- ([yshift = 0.25cm]s.west);

    \node at (7,-1) (f) {Frontier};
    \draw [->  ] (s.east) -- (f.west);

    \draw [->  ,rounded corners] (t.south) -- ([yshift = -0.5cm]t.south) -- ([yshift = -0.5cm] s.south |- t.south) -- (s.south);
    \node at ([xshift = 0.5cm,yshift = -0.75cm]s.south) {Spec};

    \node at (4,-3.5) {\textbf{\textsc{Wake: Problem Solving}}};

  \end{tikzpicture}

  \vspace{1cm}

  \begin{tikzpicture}
    \node at (0,0) (d){DSL};
    \node at ([xshift = 3cm]d.east) (p2){program};
    \node at ([yshift = 1.5cm]p2) (p1){program};
    \node at ([yshift = -1.5cm]p2) (p3){program};

    \draw[squiggle,-> ] (d.east) -- node[above]{\small sample} (p2.west);
    \draw[squiggle,-> ] (d.east) -- (p1.west);
    \draw[squiggle,-> ] (d.east) -- (p3.west);

    \node at ([xshift = 2cm]p1.east) (t1){task};
    \node at ([xshift = 2cm]p2.east) (t2){task};
    \node at ([xshift = 2cm]p3.east) (t3){task};
    \draw [-> ] (p1.east) -- node[above]{\small execute} (t1.west);
    \draw [-> ] (p3.east) -- node[above]{\small execute} (t3.west);
    \draw [-> ] (p2.east) -- node[above]{\small execute} (t2.west);

    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Dreaming}}};
  \end{tikzpicture}

  \begin{tikzpicture}
    \node at (0,0) (f1){Frontier$_1$};
    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
    \node at ([yshift = -1cm]f2.south) (f3){Frontier$_3$};

    \node at ([xshift = 2cm]f1.east) (p1){program$_1$};
    \node at ([xshift = 2cm]f2.east) (p2){program$_2$};
    \node at ([xshift = 2cm]f3.east) (p3){program$_3$};

    \draw [->,squiggle ] (f1.east) -- node[above]{\small sample} (p1.west);
    \draw [->,squiggle ] (f2.east) -- node[above]{\small sample} (p2.west);
    \draw [->,squiggle ] (f3.east) -- node[above]{\small sample} (p3.west);

    \node at ([xshift = 1.5cm]p1.east) (t1){task$_1$};
    \node at ([xshift = 1.5cm]p2.east) (t2){task$_2$};
    \node at ([xshift = 1.5cm]p3.east) (t3){task$_3$};

    \node at ([yshift = -0.5cm]p3.south) {\textsc{\textbf{Sleep-R: Experience Replay}}};
  \end{tikzpicture}
  \vspace{2cm}

  \begin{tikzpicture}
    \node at (0,0) (f1){Frontier$_1$};
    \node at ([yshift = -1cm]f1.south) (f2){Frontier$_2$};
    \node at ([yshift = -1cm]f2.south) (ff){$\vdots$};
    \node at ([yshift = -1cm]ff.south) (f3){Frontier$_N$};

    \node(c)[rectangle, rounded corners, draw, minimum width = 3cm, minimum height = 6cm, anchor = north west] at (2,1) {};
    \node[anchor=north] at (c.north) {Compression};

    \draw [-> ] (f1.east) -- (c.west|-f1.east);
    \draw [-> ] (f2.east) -- (c.west|-f2.east);
    \draw [-> ] (f3.east) -- (c.west|-f3.east);

    \node[align = right](d) at ([xshift = 1.5cm,yshift = 1cm]c.east) {DSL $\mathcal{D}$};
    \node[align = right](t) at ([xshift = 1.5cm,yshift = -1cm]c.east) {Parameters $\theta$};
    \draw [-> ] (c.east) -- (d.west);
    \draw [-> ] (c.east) -- (t.west);

    \node at (c.center) {
\begin{tikzpicture}[scale=0.7]
    %% \node[rotate=30] at (-2,0) {\begin{tabular}{c}
    %%     \footnotesize Program:\\
    %%     \code{($\lambda$ (x) (+ (- x) 1))}
    %% \end{tabular}};
    %\node at (,0.5) {\code{cons}};
%    \node [rotate=90] at (-2.3,-0.5) {\small program};

  \begin{scope}[shift={(-1,-2.5)}]
      \node(l1) at (0,0) {};
  \node[color=pop3](p1) at (-1,-1) {\code{+}};
  \node[color=pop3](n1) at (0.7,-0.9) {\code{1}};
  %\node(x1) at (0,-1) {};
  \draw[color=pop3] (l1.south) -- (p1.north);
  \draw[color=pop3] (l1.south) -- (n1.north);
  \draw[color=pop3] (-0.5,-0.45) -- (0,-1);

%  \node [rotate=90] at (-2.3,-0.7) {\small program};

\begin{scope}[shift={(0,-5)}]
  \node[pop3](p1) at (-1,-1) {\code{+}};
  \node[pop3](n1) at (0.8,-0.7) {\code{1}};
  \node[pop3](a) at (0,-1) {\code{ }};
  %\node(x1) at (0,-1) {};
  \draw[pop3] (0,0) -- (p1.north);
  \draw[pop3] (0,0) -- (n1.north);
  \draw[pop3] (-0.55,-0.4) -- (a.north);
%  \node [rotate=90] at (-2.3,-0.7) {\small fragment};

\end{tikzpicture}
    };

    \node at ([yshift=-2.5cm,xshift = 4cm]c.south) {\textsc{\textbf{Sleep-G: Memory Consolidation}}};

\begin{figure}[h]
\centering
%  \includegraphics{figures/iterations.eps}
\begin{tikzpicture}
  \begin{scope}[shift={(2.5,0)}]
    \node[obs] at (3,3) (dx){$\mathcal{D}$};
    \node[latent] at (3.5,1.75) (zp){$p$};
%    \node[obs] at (4,3) (tx){$\theta^{(x)}$};
    \node[obs] at (3.5,0.7) (xp) {$x$};
    \node[align = center] at ([yshift = -0.6cm,xshift = 0.5cm]xp.south) {Wake: Infer $p$};
    \draw [->,red] (xp.east) to[out = 0,in = 90] node(nn){} (zp.north);
%    \draw [->,red] (tx) -- (zp);
    \draw [->] (dx) -- (zp);
    \node at (nn) {
      \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.2,shift={+(-1,2.5)}}]
        \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
        \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
        \node[rectangle] at (1,1) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (I-\name) at (0,-\y) {};
        \foreach \name / \y in {1,...,3}
            \node[neuron] (H-\name) at (1,-\y-0.5) {};
        \foreach \name / \y in {1,...,4}
            \node[neuron] (O-\name) at (2,-\y) {};
        \foreach \source in {1,...,4}
            \foreach \dest in {1,...,3}
                \draw [-latex] (I-\source) -- (H-\dest);
        \foreach \source in {1,...,3}
            \foreach \dest in {1,...,4}
                \draw [-latex] (H-\source) -- (O-\dest);
      \end{tikzpicture}
    };
    \node[shift={+(0,-0.65)}] at (nn) [fill=white]{ $q(x)$ };
  \end{scope}

### 2018-07-05 poster.tex
algorithm flowchart

\usepackage{tikz}
\usetikzlibrary{arrows,shapes}

\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{tabularx}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}

@@

  \textbf{Contribution:} the \system algorithm,
  which bootstraps a learned DSL while jointly training
  a neural net to search for programs in the learned DSL

  \textbf{Approach:} given a few hundred programming tasks, alternatingly:
  \begin{itemize}
  \item \textbf{Wake}: synthesize programs
  \item \textbf{Sleep-R}: train neural net (\textbf{R}ecognition model)
    \item \textbf{Sleep-G}: improve DSL (\textbf{G}enerative model)
    \end{itemize}

  \textbf{Program representation:}
  Lisp-like; conditionals, variables, local functions

@@

@@

\begin{block}{Fragment Grammars: Inducing a DSL}
Fragment grammars: introduced in computational linguistics~(O'Donnell~2015)
\begin{figure}[b!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{7.5cm}
  \begin{tikzpicture}[scale=2]
    \node [rotate=90] at (-2.5,-0.5) {prog.};

@@

\begin{block}{Why this works: Bootstrapping}
\begin{figure}\centering
  \begin{tikzpicture}[scale=3]
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};

    \draw [->,thick] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
    \draw [->,thick] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
    \draw [->,thick] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);

    \draw [->,thick,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,thick,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
 % \caption{\system solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
\end{figure}

>>>>

\usepackage{tikz}
\usetikzlibrary{arrows.meta}
%\usetikzlibrary{arrows,shapes}
\usetikzlibrary{fit,bayesnet}
\usepackage{booktabs} % Top and bottom rules for tables
\usepackage{tabularx}
%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------
\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}

@@

  \textbf{Approach:}  \system algorithm,
  which bootstraps a learned DSL while jointly training
  a neural net to search for programs in the learned DSL.
Given a few hundred programming tasks, alternatingly:
  \begin{itemize}
  \item \textbf{Wake}: synthesize programs
  \item \textbf{Sleep-R}: train neural net (\textbf{R}ecognition model)
    \item \textbf{Sleep-G}: improve DSL (\textbf{G}enerative model)
    \end{itemize}

  \textbf{Program representation:}
  $\approx $Lisp; conditionals, variables, $\lambda$~abstraction

@@

\begin{tikzpicture}[scale=4,line width=1.5mm]

  \node[latent,scale=3] at (3.5,3) (dx){$\mathcal{D}$};
  \node[latent,scale=3] at ([xshift=2cm]dx) (zp){$p_n$};
  \node[obs,scale=3] at ([xshift=2cm]zp) (xp) {$x_n$};
  \plate {}{(zp)(xp)}{};
%  \node[align = center] at ([yshift = -1.2cm,xshift = 1cm]xp.east) {(b) \\recognition model};
  \draw [->,red] (xp.south) to[out = -90,in = -90] node(nn){} (zp.south);
%  \draw [->,red] (tx.west) -- (zp.east);
  \draw [->] (dx.east) -- (zp.west);
  \draw [->] (zp.east) -- (xp.west);

  \node at (nn) {
    \begin{tikzpicture}[x=2.5cm,y=1.25cm,transform canvas={scale=0.5,shift={+(-1,2.5)}}]
      \tikzstyle{neuron}=[circle,fill=blue!50,minimum size=20pt]
      \fill[fill=white] (-0.25,-0.5) rectangle (2.25,-4.5);
      \node[rectangle] at (1,1) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (I-\name) at (0,-\y) {};
      \foreach \name / \y in {1,...,3}
          \node[neuron] (H-\name) at (1,-\y-0.5) {};
      \foreach \name / \y in {1,...,4}
          \node[neuron] (O-\name) at (2,-\y) {};
      \foreach \source in {1,...,4}
          \foreach \dest in {1,...,3}
              \draw [-latex] (I-\source) -- (H-\dest);
      \foreach \source in {1,...,3}
          \foreach \dest in {1,...,4}
              \draw [-latex] (H-\source) -- (O-\dest);
    \end{tikzpicture}
  };
  \node[shift={+(0,-1.7)}] at (nn) { $Q$  };

  \end{tikzpicture}

@@

\begin{block}{Fragment Grammars: Inducing a DSL}
Fragment grammars: introduced in computational linguistics~(O'Donnell~2015)
\begin{figure}[b!]\tabcolsep=2pt
\centering  \begin{minipage}[c]{7.5cm}
  \begin{tikzpicture}[scale=2,line width=1mm]
    \node [rotate=90] at (-2.5,-0.5) {prog.};

@@

\begin{block}{Why this works: Bootstrapping}

\begin{figure}
\hspace*{-1.5cm}  \begin{tikzpicture}[scale=3,line width=1.5mm]
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};

    \draw [->] (synthesis.-120) to[out = -150,in = 60] node[below,rotate = 45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-G)}} (DSL.30);
    \draw [->] (synthesis.-60) to[out = -30,in = 120] node[below,rotate=-45,align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.150);
    \draw [->] (DSL.east) to[out = -30,in = 210] node[above, align = center]{{\footnotesize Trains}\\{\footnotesize (Sleep-R)}} (recognitionModel.west);

    \draw [->,dashed] (DSL.north) to[out = 90,in = 180] node[fill=white,align = center]{  \footnotesize{Inductive bias}\\\footnotesize{(Wake)}} (synthesis.west);
    \draw [->,dashed] (recognitionModel.north) to[out = 90,in = 0] node[fill=white,align = center]{{\footnotesize Makes tractable}\\{\footnotesize (Wake)}} (synthesis.east);
  \end{scope}
    \end{tikzpicture}
 % \caption{\system solves for programs, the DSL, and a recognition model. Each of these steps bootstrap off of the others in a Helmholtz-machine inspired wake/sleep inference algorithm.}  \label{feeding}
\end{figure}
\begin{itemize}
\item Search finds new programs$\implies$ DSL+recognition model get more data
\item DSL improves$\implies$easier search, recognition model gets better data
  \item Recognition model improves$\implies$ easier search
  \end{itemize}
